---
title: "Time series analysis with 'polmineR'"
subtitle: 'Sample analyses with the GermaParl corpus'
author: "Andreas Blaette"
date: "13 Mai 2018"
output:
  slidy_presentation:
    footer: Copyright (c) 2018, Andreas Blaette
  beamer_presentation: default
  ioslides_presentation: default
editor_options: 
  chunk_output_type: console
---


## Initialisiation

The examples are based on the GermaParl corpus. The data set in the package must be activated with the `use` command after loading polmineR.

```{r initialize, eval = TRUE, message=FALSE}
library(polmineR)
use("GermaParl")
```

If you want to follow the code of the following examples on your own computer, you can load the 'Rmarkdown' document, with which the slide set was generated, directly into R:

```{r get_code, eval = FALSE}
if (!"RCurl" %in% rownames(installed.packages())) install.packages("RCurl")
library(RCurl)
script <- getURL(
  "https://raw.githubusercontent.com/PolMine/polmineR.tutorials/master/polmineR_Zeitreihen.Rmd",
  ssl.verifypeer = FALSE
  )
```

Save the script to your desktop - or to another location.

```{r save_code, eval = FALSE}
filename <- "~/Desktop/polmineR_Zeitreihen.Rmd"
if (interactive()) writeLines(text = script, con = filename)
```

If you now open the document, you can follow the code step by step (and change it according to your requirements)

----

## Time and frequency distributions

The frequency of words (in the sense of keywords) or search terms can be determined with the `count` method. The `dispersion` method performs the count using one or two structural attributes (i.e. dimensions).

```{r get_simple_dispersion}
dt <- dispersion("GERMAPARL", query = "Flüchtlinge", sAttribute = "year")
head(dt) # we consider only the beginning of the table
```

A nice and practical view is offered by the `View` function of RStudio.

```{r, eval = interactive()}
View(dt)
```

The return value of the `dispersion` method is a `data.table`. If you are used to working with `data.frame` objects, you can easily convert it.

```{r}
df <- as.data.frame(dt)
```

----

## Simple visualization of frequencies

```{r}
barplot(height = df[["count"]], names.arg = df[["year"]], las = 2)
```

This simple bar chart already confirms the assumption that the sharp increase in forced migration in 2015 also had a considerable impact on the German Bundestag.

----

## Frequency distribution in two dimensions

We now add a second dimension to the analysis: differentiation by political parties.

```{r}
dt <- dispersion("GERMAPARL", query = "Flüchtlinge", sAttribute = c("year", "party"))
```

To work with time series data, we use the `xts` package, which will be installed first if required.

```{r, message = FALSE}
if (!"xts" %in% rownames(installed.packages())) install.packages("xts")
library(xts)
```

We create an `xts` object based on the already existing crosstable containing the frequencies and see what it looks like.

```{r}
ts <- xts(x = dt[,c("CDU", "CSU", "FDP", "GRUENE", "SPD")],
          order.by = as.Date(sprintf("%s-01-01", dt[["year"]])))
head(ts)
```

----

## Visualization of time progression

Things are better seen through a time series diagram.

```{r}
plot(ts, multi.panel = TRUE, col = c("black", "black", "blue", "green", "red"), lwd = 2)
```

----

## What happens throughout the years?

```{r}
dt <- dispersion("GERMAPARL", query = "Flüchtlings", sAttribute = "date")
ts <- xts(x = dt[["count"]], order.by = as.Date(dt[["date"]]))
plot(ts)
```

That doesn't look very convincing. We should therefore find a way to aggregate the census over a longer time period.

---- 

## Week - month - quarter - year

We will use the week, month, quarter and year as time units for aggregation beyond the individual day. For the weeks we need the `lubridate` package.

```{r load_lubridate, message = FALSE}
if (!"lubridate" %in% rownames(installed.packages())) install.packages("lubridate")
library(lubridate)
```

Now we create aggregated time series objects. The code for this is intentionally compact and perhaps not immediately comprehensible. In case of doubt ... use copy & paste!

```{r}
ts_week <- aggregate(ts, {a <- ymd(paste(year(index(ts)), 1, 1, sep = "-")); week(a) <- week(index(ts)); a})
ts_month <- aggregate(ts, as.Date(as.yearmon(index(ts))))
ts_qtr <- aggregate(ts, as.Date(as.yearqtr(index(ts))))
ts_year <- aggregate(ts, as.Date(sprintf("%s-01-01", gsub("^(\\d{4})-.*?$", "\\1", index(ts)))))
```

----

Which time series aggregation makes the most sense to you?

```{r}
par(mfrow = c(2,2), mar = c(2,2,3,1))
plot(ts_week, main = "Aggregation: Week"); plot(ts_month, main = "Aggregation: Month");
plot(ts_qtr, main = "Aggregation: Quarter"); plot(ts_year, main = "Aggregation: Year")
```
