---
title: "Usage of concordances in 'polmineR'"
subtitle: 'Sample analyses with the GermaParl corpus'
author: "Andreas Blaette"
date: "28 Mai 2018"
output:
  slidy_presentation:
    footer: Copyright (c) 2018, Andreas Blaette
  beamer_presentation: default
  ioslides_presentation: default
editor_options: 
  chunk_output_type: console
---


```{r, message = FALSE, results = "hide", echo = FALSE}
# This code is not displayed in the slide set and is only required,
# to reload any missing packages:
# kable and kableExtra are needed to generate tabular output.
if (!"knitr" %in% rownames(installed.packages())) install.packages("knitr")
if (!"kableExtra" %in% rownames(installed.packages())) install.packages("kableExtra")
```


## Initialisiation

The examples are based on the GermaParl corpus. The data set in the package must be activated with the `use` command after loading polmineR.

```{r initialize, eval = TRUE, message=FALSE}
library(polmineR)
use("GermaParl")
```

If you want to follow the code of the following examples on your own computer, you can load the 'Rmarkdown' document, with which the slide set was generated, directly into R:

```{r get_code, eval = FALSE}
if (!"RCurl" %in% rownames(installed.packages())) install.packages("RCurl")
library(RCurl)
script <- getURL(
  "https://raw.githubusercontent.com/PolMine/polmineR.tutorials/master/polmineR_KWIC.Rmd",
  ssl.verifypeer = FALSE
  )
```


Save the script to your desktop - or to another location.

```{r save_code, eval = FALSE}
filename <- "~/Desktop/polmineR_KWIC.Rmd"
if (interactive()) writeLines(text = script, con = filename)
```

If you now open the document, you can follow the code step by step. Use the code as a starting point for your own analyses!

----

## Concordances and KWIC: Basics

  * The analysis of word surroundings of terms and search strings serves as an analytical bridge between quantitative counting methods and a qualitative-interpretative approach to text data. Linguistics calls this concordances, while the social science tradition of content analysis refers to keyword-in-context analyses ("KWIC" for short). For brevity and clarity, the method in the polmineR package is called `kwic`.
  
  * An important preliminary decision is to define how many words are to be displayed to the left and to the right. In linguistic analyses (expecially in lexicography, i.e. development of lexica) a window of 5 words is used left and right. How much context do you as a political scientist need for a rough contextualization? Often this will be more than in the linguistic usage of concordances / the KWIC procedure.
  
  * In some cases it will not be sufficient to read and interpret only a short text excerpt in the context of a word. If necessary, the full text of a search hit should be read. This is achieved in the polmineR package by the `read` method.
  

----

## Concordances and KWIC: During the research process

  * The analysis of concordances can also be usefully combined with the statistical analysis of coccurrences: The coocurrency analysis provides information on static conspicuities of language usage, which are then analysed and interpreted in more detail using KWIC.
  
  * A useful step for systematizing interpretations of concordances is to categorize or classify them. It can be helpful to export them to Excel (via the `mail` method) in order to enter categorizations or note down interpretations in an additional column there.
  
  * When working with concordances, please be aware that this is an interpretative work that requires a hermeneutic feeling!

----

## KWIC method: The basics

As with other functions and methods in R, you can call the documentation for the `kwic` method by placing a question mark in front of it. You will also find first bibliographical references there!

```{r, eval = FALSE}
?kwic
```

The method can be applied to objects of the classes `character` (i.e. to a corpus as a whole), `partition` and `context`. You define the search term using the `query` argument.

```{r, eval = FALSE}
kwic("GERMAPARL", query = "Migrationshintergrund")

bt2005 <- partition("GERMAPARL", year = 2005)
kwic(bt2005, query = "Migrationshintergrund")
```  

When formulating search queries, you can use the same CQP syntax. Set the argument `cqp` to `TRUE`. (If you do not do this, polmineR automatically checks whether the CQP syntax is used.)

```{r, eval = FALSE}
kwic("GERMAPARL", query = '[pos = "NN"] "mit" "Migrationshintergrund"', cqp = TRUE)
kwic(bt2005, query = '[pos = "NN"] "mit" Migrationshintergrund"', cqp = TRUE)
```  

The results are then displayed in RStudio's Viewer Panel.

----

## KWIC: Application scenarios

As discussed earlier, you will often want to see more of the word surroundings than just (as default) five words left and right. Use the arguments `left` and `right` to determine the word range.

```{r, eval = FALSE}
kwic("GERMAPARL", query = "Ausländer", left = 10, right = 10)
kwic("GERMAPARL", query = "Ausländer", left = 15, right = 15)
```

There are good reasons to assume that characteristics of a speaker such as parliamentary group or party affiliation (thesis of party differences!), gender etc. make a difference: Display the metadata for a concordance with the `sAttribute` argument. Note: Several attributes must be combined using c(), the "c" here stands for "combine".

```{r, eval = FALSE}
kwic(bt2005, query = "Ausländer", sAttribute = "party")
kwic(bt2005, query = "Ausländer", sAttribute = c("party", "date"))
```

To read the full text of a concordance, perform a temporary save of the KWIC result, and then apply the `read` method.

```{r, eval = FALSE}
K <- kwic(bt2005, query = "Ausländer", sAttribute = "party")
read(K, 1)
```
