---
title: "Using the CQP syntax with 'polmineR'"
subtitle: 'Sample analyses with the GermaParl corpus'
author: "Andreas Blaette"
date: "28 Mai 2018"
output:
  slidy_presentation:
    footer: Copyright (c) 2018, Andreas Blaette
  beamer_presentation: default
  ioslides_presentation: default
editor_options: 
  chunk_output_type: console
---


```{r, message = FALSE, results = "hide", echo = FALSE}
# This code is not displayed in the slide set and is only required,
# to reload any missing packages:
# kable and kableExtra are needed to generate tabular output.
if (!"knitr" %in% rownames(installed.packages())) install.packages("knitr")
if (!"kableExtra" %in% rownames(installed.packages())) install.packages("kableExtra")
```


## Initialisation

The examples are based on the GermaParl corpus. The data set in the package must be activated with the `use` command after loading polmineR.

```{r initialize, eval = TRUE, message=FALSE}
library(polmineR)
use("GermaParl")
```

If you want to follow the code of the following examples on your own computer, you can load the 'Rmarkdown' document, with which the slide set was generated, directly into R:

```{r get_code, eval = FALSE}
if (!"RCurl" %in% rownames(installed.packages())) install.packages("RCurl")
library(RCurl)
script <- getURL(
  "https://raw.githubusercontent.com/PolMine/polmineR.tutorials/master/polmineR_CQP.Rmd",
  ssl.verifypeer = FALSE
  )
```

Save the script to your desktop - or to another location.

```{r save_code, eval = FALSE}
filename <- "~/Desktop/polmineR_CQP.Rmd"
if (interactive()) writeLines(text = script, con = filename)
```

If you now open the document, you can follow the code step by step. Use the code as a starting point for your own analyses!
----

## CWB, CQP and 'polmineR'

The R package uses the so-called *Corpus Workbench* (CWB) to store and retrieve data. The CWB is a bundle of tools with the functionality of an "indexing and query engine". Comparable software that offers advantages for industrial applications would be [Lucene](https://lucene.apache.org/) or [Elasticsearch](https:://www.elastic.co). However, a major advantage of the CWB is the fact that it is completely open source and can be used and further developed in software projects.

EA powerful part of the CWB is the so-called *Corpus Query Processor* (CQP) as an instrument for formulating complex linguistic search queries for large corpora, which can contain regular expressions and linguistic annotations of corpora. The CQP syntax can be used for all basic commands of the polmineR package:

  * `kwic` (Keyword-in-Context analysis / query of concordances);
  * `count` (Frequency counting in corpora and sub-corpora);
  * `dispersion` (Distribution of search hits using a structural attribute);
  * `cooccurrences` (Analysis of coccurrences in the surrounding word of a search hit);

In the following examples, the `count` method is used to show how the CQP syntax works. Remember that the other functions are also capable of CQP! The following tutorial briefly introduces the use of regular expressions. We will then explain the use of linguistic features of a corpus.


## Search for words

The CQP search syntax is passed to the methods with the parameter `query`. The default setting is automatic detection of whether the CQP syntax is used. However, it is recommended to specify explicitly that CQP is used using the logical parameter `cqp`.

It is important that individual search terms must be enclosed in quotation marks and single quotation marks enclose the entire query.

```{r}
count("GERMAPARL", query = '"Diskriminierung"', cqp = TRUE)
```

The *flag* %c can be used to remove the distinction between upper and lower case. This leads, for example, to large differences in the counting of all regular lowercase words (adjectives, verbs, etc.), because they are of course capitalized at the beginning of a sentence.

```{r}
count("GERMAPARL", query = c('"Liebe"', '"Liebe" %c'), cqp = TRUE)
```


----

## Regular expressions: character classes

A key feature of the CQP syntax is that when searching for individual words, the entire set of regular expressions can be used. Regular expressions are a standardized tool available in virtually all programming languages for searching for character patterns in text.

| Expression | Description |
|:-------:| --------------|
| .       |A dot (".") stands for any character |
| \\d | "digit", e.g. 0 to 9 |
|?|The preceding expression does not occur or occurs once.|
|+|The preceding expression occurs once or more than once. |
|*|The preceding expression does not occur at all or can occur at will.|
|{n}|The preceding expression occurs exactly n times.|
|{min,}| The preceding expression occurs at least min times.|
|{min,max}|The preceding expression occurs at least min times and occurs maximally max times.|
|{0,max}| The preceding expression may occur max times.|

----

## Regular expressions: Examples I

This makes it possible to design simple searches:

```{r}
y <- count("GERMAPARL", query = '"Multikult.*"', cqp = TRUE, breakdown = TRUE)
head(y, n = 3)
```

It is important that alternative characters are enclosed in square brackets. If we don't want to work with the *flag* %c we can also search for ...

```{r}
y <- count("GERMAPARL", query = '"[mM]ultikult.*"', cqp = TRUE, breakdown = TRUE)
head(y, n = 3)
```


## Regular expressions: Examples II

While alternative individual letters may be specified in square brackets, alternative character strings may be specified in round brackets, each separated by a vertical dash ("|").


```{r}
y <- count("GERMAPARL", query = '"(Zu|Ein|Aus)wanderung.*"', breakdown = TRUE)
head(y)
```

```{r}
y <- count("GERMAPARL", query = '"Asyl(suchende|berechtigte|ant|anti)"', cqp = TRUE, breakdown = TRUE)
head(y)
```

----


## CQP I: Basics

CQP extends the syntax of regular expressions to include options to formulate queries using multiple words ("tokens"). CQP can query the different annotations of a linguistically annotated corpus. The data format is always in tabular form. This means that if, in addition to the original word form (p attribute "word"), a word type recognition (so-called "part of speech" notation, p attribute "pos") and lemmatization (p attribute "lemma") has been executed, a three-column *tokenstream* is available, which can look as follows at the beginning of a plenary debate, for example. The first column shows the *corpus position* ("cpos"):

```{r, echo = FALSE, message = FALSE}
P <- partition("GERMAPARL", speaker = "Angela Merkel", lp = "15")
cpos_left <- P@cpos[1,1]
pAttributes <- c("word", "pos", "lemma")
tokenstream_list <- lapply(
  pAttributes,
  function(x) getTokenStream("GERMAPARL", pAttribute = x, encoding = "latin1", left = cpos_left, right = cpos_left + 9)
)
tokenstream_df <- as.data.frame(tokenstream_list)
colnames(tokenstream_df) <- pAttributes
tokenstream_df[["pos"]] <- gsub("^\\$", "\\\\$", tokenstream_df[["pos"]])
tokenstream_df[["cpos"]] <- 0L:9L
tokenstream_df <- tokenstream_df[, c("cpos", pAttributes)]
K <- knitr::kable(tokenstream_df, format = "html", escape = TRUE)
kableExtra::kable_styling(K, font_size = 18, position = "left")
```

---- 

## CQP II: Searching in the tokenstream

If you use the CQP syntax to search in the p attribute "word", the search term must only be enclosed in quotation marks. To access the other p attributes, the attribute to which you want to refer is specified in square brackets. With the following search term "Q" you search for sequences of a noun, then "with" and a word that begins with "Migrations".


```{r, eval = TRUE}
use("GermaParl")
Q <- '[pos = "NN"] "mit" "Migrations.*"'
C <- count("GERMAPARL", query = Q, breakdown = TRUE)
head(C[,c("match", "count", "share")])
```

----

## CQP III: Quantizers

Empty square brackets serve as placeholders for any words. For example ... 

```{r}
y <- count("GERMAPARL", query = '"(Bundesm|M)inisterium" [] [pos = "NN"]', cqp = TRUE, breakdown = TRUE)
head(y[,c("match", "count", "share")], n = 3)
```

According to the notation already known from the regular expressions, you can specify in curved brackets how many arbitrary words should occur.

```{r}
y <- count("GERMAPARL", query = '"([Kk]riminell.*|Straftat.*)" []{0,5} "Asyl.*"', cqp = TRUE, breakdown = TRUE)
head(y[,c("match", "count", "share")], n = 3)
```


---- 

## CQP: All polmineR methods can use this

As mentioned above, the CQP syntax is available for all basic methods of the polmineR package.

```{r}
K <- kwic("GERMAPARL", query = '"Integration" []{0,5} ".*[Ss]cheiter.*"', cqp = TRUE)
```

```{r}
D <- dispersion("GERMAPARL", query = '"[rR]assis.*"', sAttribute = "party")
```

```{r}
C <- cooccurrences("GERMAPARL", query = '"([mM]uslim.|[iI]slam*)"', cqp = TRUE)
```

Of course, this is also the case if you apply these basic methods to partitions.

```{r}
post2001 <- partition("GERMAPARL", year = 2002:2009, verbose = FALSE)
C <- cooccurrences(post2001, query = '"([mM]uslim.|[iI]slam*)"', cqp = TRUE)
```

CQP is a powerful analysis tool! Formulating the right search queries, however, requires some practice. Pay particular attention to this:

*  Don't forget to put CQP queries in single opening and closing quotes!
  
* If you receive an error message, check whether opening quotation marks, square brackets or curved brackets are closed!

Good luck! 



----

## Appendix: The Stuttgart-Tübingen Tagset (excerpt)

|Notation|Description|Example|
|:------:|------------|--------|
|ADJA|attributives Adjektiv| [das] große [Haus]|
|ART| bestimmter oder unbestimmter Artikel | der, die, das, ein, eine, ... |
|NN| normales Nomen|Tisch, Herr, [das] Reisen|
|NE|Eigennamen| Hans, Hamburg, HSV|
|VVFIN|finites Verb, voll | [du] gehst, [wir] kommen [an] |
|VVIMP|Imperativ, voll|komm [!]|
|VVINF|Infinitiv, voll|gehen, ankommen|
|VVIZU|Infinitiv mit ``zu'', voll| anzukommen, loszulassen|
|VVPP|Partizip Perfekt, voll| gegangen, angekommen|
|VAFIN|finites Verb, aux| [du] bist, [wir] werden | 



